Global adjoint tomography -- Model GLAD-M25\footnote{This paper is dedicated to Dimitri Komatitsch, computational geophysicist extraordinaire, who passed away unexpectedly shortly before its submission.}

### 2019-01-30 main.tex
Updates from ShareLaTeX

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
Building on our first-generation global adjoint tomography model,
GLAD-M15, we present our second-generation transversely isotropic global model,
named GLAD-M25, which is the result of ten quasi-Newton tomographic iterations with an
earthquake database consisting of 1,480 events in the magnitude range $5.8\le M_w \le 7.0$,
an almost six-fold increase over the first-generation model.
We calculated fully 3D synthetic seismograms with a shortest period of 17~s based on a GPU-accelerated spectral-element wave
propagation solver which accommodates effects due to 3-D anelastic crust \& mantle structure, topography \& bathymetry, the ocean load, ellipticity, rotation, and self-gravitation.
We used an adjoint-state method to calculate Fr\'echet derivatives in 3D anelastic Earth models
facilitated by a parsimonious storage algorithm.
The simulations were performed on the Cray XK7 ``Titan'' at the Oak Ridge Leadership Computing Facility.
We quantitatively evaluated GLAD-M25 by assessing misfit reductions and traveltime \& amplitude anomaly histograms in twelve measurement
categories.
We performed similar assessments for a held-out data set consisting of 380 earthquakes,
with results comparable to the actual inversion.
We compare GLAD-M25 with numerous other global and regional models,
and highlight a variety of plumes and subduction zones.
\end{summary}

@@

The observed seismograms need to be compared to corresponding synthetic seismograms
for the purpose of making measurements.
The forward simulations are based on the global spectral-element solver
SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02a}, and take a CMT solution, a relevant list
of stations, and the current earth model as input.
The solver is GPU accelerated, and the simulations are performed on the Cray supercomputer ``Titan''
at the OLCF.
To simulate 120~min seismograms with a shortest period of 17~s on 384 NVIDIA
Tesla K20X GPUs takes approximately 10~min of wall clock time.
The calculation of Fr\'echet derivatives at a later stage of the workflow requires
access to snapshots of the forward wavefield as part of the parsimonious storage
algorithm developed by~\cite{KoXiBoPeSaLiTr16},
with rapid I/O facilitated by ADIOS~\citep{liu2014hello}.
For the current simulation setup,
this requires about 1~TB of storage per earthquake,
so for 1,480 earthquakes this amounts to 1.5~PB of storage.

>>>>

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
Building on our first-generation global adjoint tomography model,
GLAD-M15, we present our second-generation transversely isotropic global model,
named GLAD-M25, which is the result of ten quasi-Newton tomographic iterations with an
earthquake database consisting of 1,480 events in the magnitude range $5.8\le M_w \le 7.0$,
an almost six-fold increase over the first-generation model.
We calculated fully 3D synthetic seismograms with a shortest period of 17~s based on a GPU-accelerated spectral-element wave
propagation solver which accommodates effects due to 3-D anelastic crust \& mantle structure, topography \& bathymetry, the ocean load, ellipticity, rotation, and self-gravitation.
We used an adjoint-state method to calculate Fr\'echet derivatives in 3D anelastic Earth models
facilitated by a parsimonious storage algorithm.
The simulations were performed on the Cray XK7 ``Titan'' at the Oak Ridge Leadership Computing Facility.
We quantitatively evaluated GLAD-M25 by assessing misfit reductions and traveltime \& amplitude anomaly histograms in twelve measurement
categories.
We performed similar assessments for a held-out data set consisting of 360 earthquakes,
with results comparable to the actual inversion.
We compare GLAD-M25 with numerous other global and regional models,
and highlight a variety of plumes and subduction zones.
\end{summary}

@@

The observed seismograms need to be compared to corresponding synthetic seismograms
for the purpose of making measurements.
The forward simulations are based on the global spectral-element solver
SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02a}, and take a CMT solution, a relevant list
of stations, and the current earth model as input.
The solver is GPU accelerated, and the simulations are performed on the Cray supercomputer ``Titan''
at the OLCF.
To simulate 120~min seismograms with a shortest period of 17~s on 384 NVIDIA
Tesla K20X GPUs takes approximately 10~min of wall clock time.
The calculation of Fr\'echet derivatives at a later stage of the workflow requires
access to snapshots of the forward wavefield as part of the parsimonious storage
algorithm developed by~\cite{KoXiBoPeSaLiTr16},
with rapid I/O facilitated by ADIOS~\citep{liu2014hello}.
For the current simulation setup,
this requires about 1~TB of storage per earthquake,
so for 1,480 earthquakes this adds up to 1.5~PB.
This volume of data is generated during~$\sim6$~hours of simulations.
Based on our experience, ADIOS is capable of achieving~$\sim120$~GB/s
of sustained I/O on OLCF resources.
Given that typical write speeds for modern hard disks are~$\sim100$~MB/s,
it would be extremely difficult to carry out global adjoint tomography without the ADIOS library.
